---
title: 'Class 8: Genome Informatics'
author: "Katherine Wong (A16162648)"
date: "12/4/2021"
output: 
  github_document:
  pdf_document:
  html_document: default
---
## Section 4: Population Scale Analysis

One sample is obviously not enough to know what is happening in a population. You are interested in assessing genetic differences on a population scale. 

So, you processed about ~230 samples and did the normalization on a genome level. Now, you want to find whether there is any association of the 4 asthma-associated SNPs (rs8067378...) on ORMDL3
expression.

> Q13: Read this file into R and determine the sample size for each genotype and their corresponding median expression levels for each of these genotypes.

```{r}
expr <- read.table("rs8067378_ENSG00000172057.6.txt")
head(expr)
```
```{r}
nrow(expr)
```
```{r}
table(expr$geno)
```

Sample Size for A/A: 108, A/G= 233, G/G=121. Median for A/A = 31.24847. Median for A/G = 25.06486. Median for G/G = 20.07363. 

```{r}
median(expr[expr$geno == "A/A", "exp"])
median(expr[expr$geno == "A/G", "exp"])
median(expr[expr$geno == "G/G", "exp"])
```

```{r}
library(ggplot2)
```

Let's make a boxplot

```{r}
ggplot(expr) + aes(x=geno, y=exp, fill=geno) +
  geom_boxplot(notch=TRUE)

```


