---
title: 'Class 13: Unix for Bioinformatics'
author: "Katherine Wong (A16162648)"
date: "11/9/2021"
output:
  pdf_document: default
  html_document: default
---

> Q1. What does the star character accomplish here? Ask Barry, or your class neighbor, if you are not sure!

The star character is the wildcard. This means that if we use the command 'gunzip *.faa.gz', this will uncompress all files that end with .faa.gz. 

> Q2. How many sequences are in this mouse.1.protein.faa file? Hint: Try using grep to figure this out…

67876 protein sequences. Using "grep -c ">" mouse.1.protein.faa".

> Q3. What happens if you run the above command without the > mm-first.fa part?

Prints to stdout the first two FASTA protein sequences.

> Q4. What happens if you were to use two ‘>’ symbols (i.e. >> mm-first.fa)?

It redirects the output by appending the two protein sequences to the end of the existing mm-first.fa file

> Q5. How would you determine how many sequences are in the mm-second.fa file?

grep -c ">" mm-second.fa. There are 83 sequences in the file. 

> Q6. Note the addition of the -r option here: What is it’s purpose? Also what about the *, what is it’s purpose here? 

The purpose of -r here is that it recursively copies the files and its sub-directories. So everything in the directory should be copied over, including subdirectories and the files within those subdirectories, etc. The purpose of the * or the wildcard is that it copies all the files and directories in specified directory, in this case home/work/ from the remote machine.

```{r}
zebra <- read.csv(file = "mm-second.x.zebrafish.tsv", sep = "\t", header = FALSE, col.names =c("qseqid", "sseqid", "pident", "length", "mismatch", "gapopen", "qstart", "qend", "sstart", "send", "evalue", "bitscore"))
head(zebra)
```


```{r}
hist(zebra$bitscore, breaks = 30)
hist(zebra$pident, breaks = 30)
```

```{r}
plot(zebra$pident  * (zebra$qend - zebra$qstart), zebra$bitscore)

```
```{r}
library(ggplot2)
ggplot(b, aes(pident, bitscore)) + geom_point(alpha=0.1) 
```



