---
title: 'Class 10: Halloween Candy'
author: "Katherine Wong (A16162648)"
date: "10/28/2021"
output: html_document
---

# Get the data

Here we look at 538 Halloween candy data. We take the CSV file from their Github repo. 

```{r}
url <- "https://raw.githubusercontent.com/fivethirtyeight/data/master/candy-power-ranking/candy-data.csv"

candy <- read.csv(url, row.names=1)
head(candy, n = 5)

```

> Q1. How many different candy types are in this dataset? 

```{r}
nrow(candy)
```
85 different candy types.

> Q2. How many fruity candy types are in the dataset?

```{r}
sum(candy$fruity)
```
38 fruity candy types.

> Q3. What is your favorite candy in the dataset and what is it’s winpercent value?

```{r}

candy["Twix", ]$winpercent
```
My favorite candy is Twix. The winpercent of Twix is 81.64%

> Q4. What is the winpercent value for “Kit Kat”?

```{r}

candy["Kit Kat", ]$winpercent
```
The winpercent of Kit Kat is 76.77%

> Q5. What is the winpercent value for “Tootsie Roll Snack Bars”?

```{r}

candy["Tootsie Roll Snack Bars", ]$winpercent
```
The winpercent of Tootsie Roll Snack Bars is 49.65%

```{r}
library(skimr)
skim(candy)
```
> Q6. Is there any variable/column that looks to be on a different scale to the majority of the other columns in the dataset?

winpercent

> Q7. What do you think a zero and one represent for the candy$chocolate column?

1 = chocolate
0 = not chocolate

> Q8. Plot a histogram of winpercent values

```{r}
hist(candy$winpercent)
```

> Q9. Is the distribution of winpercent values symmetrical?

no

> Q10. Is the center of the distribution above or below 50%?

below

> Q11. On average is chocolate candy higher or lower ranked than fruit candy?

First we need to find all the chocolate candy rows in the 'candy' data
```{r}
inds <- as.logical(candy$chocolate)
chocolate <- candy[inds,]$winpercent

inds.fruit <- as.logical(candy$fruity)
fruity <- candy[inds.fruit,]$winpercent

mean(chocolate)
mean(fruity)
```

On average, chocolate candy is ranked higher than fruity. 

> Q12. Is this difference statistically significant?

```{r}

t.test(chocolate, fruity)
```
Yes, the p value is less than 0.05 so it is statistically significant.

> Q13. What are the five least liked candy types in this set?

Nik L Nip, Boston Baked Beans, Chiclets, Super Bubble and Jawbusters. 

```{r}
head(candy[order(candy$winpercent),], n=5)
```

> Q14. What are the top 5 all time favorite candy types out of this set?

Reeses Peanut Butter Cups, Reeses Miniatures, Twix, Kit Kat, and Snickers.

> Q15. Make a first barplot of candy ranking based on winpercent values.


```{r}
library(ggplot2)

ggplot(candy) + 
  aes(winpercent, rownames(candy)) +
  geom_col()
```

> Q16. This is quite ugly, use the reorder() function to get the bars sorted by winpercent?

Improve this plot by reordering by the 'winpercent' value

```{r}
ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy), winpercent)) +
  geom_col()
```

```{r}
my_cols=rep("black", nrow(candy))
my_cols[as.logical(candy$chocolate)] = "chocolate"
my_cols[as.logical(candy$bar)] = "brown"
my_cols[as.logical(candy$fruity)] = "pink"

ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent)) +
  geom_col(fill=my_cols) 
```

Now, for the first time, using this plot we can answer questions like:

> Q17. What is the worst ranked chocolate candy?

```{r}
ord <- order(candy$pricepercent, decreasing = TRUE)
head( candy[ord,c(11,12)], n=5 )
```


> Q18. What is the best ranked fruity candy?

Starburst.

# Taking a look at pricepoint

```{r}
library(ggrepel)

my_cols[as.logical(candy$fruity)] = "red"

# How about a plot of price vs win
ggplot(candy) +
  aes(winpercent, pricepercent, label=rownames(candy)) +
  geom_point(col=my_cols) + 
  geom_text_repel(col=my_cols, size=3.3, max.overlaps = 5)
```
> Q19. Which candy type is the highest ranked in terms of winpercent for the least money - i.e. offers the most bang for your buck?

```{r}
ord <- order(candy$winpercent, decreasing = TRUE)
head( candy[ord,c(11,12)], n=5 )
```

Reese's Miniatures.

> Q20. What are the top 5 most expensive candy types in the dataset and of these which is the least popular?

```{r}

ord <- order(candy$pricepercent, decreasing = TRUE)
head( candy[ord,c(11,12)], n=5 )
```

Nik L Nip, Nestle Smarties, Ring pop, Hersheys Krackel, Hersheys Milk Chocolate are the top 5 most expensive candies. Nik L Nip are the least popular at a winpercent of 22.44%.

# Make a lollipop chart of pricepercent

```{r}
ggplot(candy) +
  aes(pricepercent, reorder(rownames(candy), pricepercent)) +
  geom_segment(aes(yend = reorder(rownames(candy), pricepercent), 
                   xend = 0), col="gray40") +
    geom_point()
```

# Exploring correlation structure

```{r}
library(corrplot)
cij <- cor(candy)
corrplot(cij)
```

> Q22. Examining this plot what two variables are anti-correlated (i.e. have minus values)?

chocolate and fruity.

> Q23. Similarly, what two variables are most positively correlated?

chocolate and winpercent. or chocolate and bar. or all variables with itself (value of 1)

# 6. Principal Component Analysis

```{r}
pca <- prcomp(candy, scale=TRUE)
summary(pca)
```

```{r}
plot(pca$x[,1:2])
plot(pca$x[,1:2], col=my_cols, pch=16)


```

```{r}
# Make a new data-frame with our PCA results and candy data
my_data <- cbind(candy, pca$x[,1:3])

p <- ggplot(my_data) + 
        aes(x=PC1, y=PC2, 
            size=winpercent/100,  
            text=rownames(my_data),
            label=rownames(my_data)) +
        geom_point(col=my_cols)

p
```

```{r}
library(ggrepel)

p + geom_text_repel(size=3.3, col=my_cols, max.overlaps = 7)  + 
  theme(legend.position = "none") +
  labs(title="Halloween Candy PCA Space",
       subtitle="Colored by type: chocolate bar (dark brown), chocolate other (light brown), fruity (red), other (black)",
       caption="Data from 538")
```
```{r}
library(plotly)
ggplotly(p)

```

```{r}
par(mar=c(8,4,2,2))
barplot(pca$rotation[,1], las=2, ylab="PC1 Contribution")
```

> Q24. What original variables are picked up strongly by PC1 in the positive direction? Do these make sense to you?

fruity. yes it makes sense.


